---
title: "24Sep2015"
author: "Chris Lynum"
date: "September 24, 2015"
output: html_document
---

Split, apply, combine

```{r}
library(gapminder)
library(dplyr)
library(ggplot2)
# Clear your workspace
# rm(list= ls())
```



```{r}
str(gapminder)
# selecting the rows
filter(gapminder, country == 'Canada')

# selecting columns
# gapminder[, 'country'] <- old way
select(gapminder, country)

# only finds columns that start with whatever you put
select(gapminder, starts_with('C'))

head(select(gapminder, country))
```


```{r}
group_by(gapminder, continent)
# puts all continents together (Asias together, Europes together, etc.)

# if you want to collapse a ton of data
group_by(gapminder, continent) %>%
  summarize(mean_life = mean(lifeExp))


# if you want something to happen to every row of the data
# adds a column to data set
head(mutate(gapminder, gdp = gdpPercap * pop)) 


# don't want all informoation, just wont the new column
head(transmute(gapminder, gdp = gdpPercap * pop))
```


```{r}
head(arrange(gapminder, desc(lifeExp)))
# ascending is default
```

```{r}
gapminder2 <- tbl_df(gapminder)
str(gapminder2)
# has multiple classes instead of just data frame
# allows us to use a few different packages more easily
gapminder2
# now only shows the first 10 rows intead of listing everything
```


## Practice

### 1) For each continent, what country had the smallest population in 1952, 1972, and 2002?

So I want to separate for each continent
Need to know the smallest population in each country
for three different years
```{r}
group_by(gapminder, continent)
filter(gapminder, year == c(1952, 1972, 2002)) %>% 
  group_by(continent) %>%
  group_by(country) %>% arrange(desc(pop))


filter(gapminder2, year == 1952 | year == 1972 | year == 2002)
  
#filter(gapminder2, year == c(1952, 1972, 2002)) %>%
# arrange(desc(year))

minpop <- filter(gapminder2, year %in% c(1952, 1972, 2002)) %>%
  group_by(continent, year) %>% select(min(pop)) 
View(minpop)
```

### 2) Create a graph of the total population of each continent over time
```{r}
prob2 <- ggplot(data = gapminder2, aes(x = continent, y = pop))
prob2 + geom_bar(stat = 'identity')
```


### 3) Calculate the mean gdpPercap for each country

```{r}

```

### 4) Calculate the mean gdp for each country

```{r}
group_by(gapminder, country) %>% 
  mutate(data = gapminder, gdp = gdpPercap * pop)) %>%
  summarize(mean_gdpP = mean(gdpPercap)) %>%
  summarize(mean_gdp = mean(gdp))
```

